---
title: Investigation of Policy and Legislative Actions for the Reduction of Fatal
  Traffic Incidents in Relation to Alcohol and Socio-Economic Factors
author: "Keith Mitchell, ..."
date: "2/12/2021"
output: html_document
---


# Introduction
Fatalities related to traffic is a point of interest for policy makers at the state and federal level, which could likely be associated with various alcohol, societal, religional, and socio-economic factors in these regions. 

Traffic incidents are the number one source of deaths for persons in the age range of 1-34 and 41% of fatal crashes associate with drinking (Hingson 1996) Alcohol is known to alter driving performance, and the various factors surrounding alcohol legislation as well as other societal factors is likely to modify alcohol usage by pupils and therefore traffic fatalities (Zhao, Zhang, and Rong 2014). In addition, a variety of socio-economic factors have also been previously studied showing that traffic accidents for certain economic levels are associated with increases in mortality (Sehat et al. 2012). Legislation and policy can help to diminish the effect of alcohol related incidences as well as lower the effect of social and economic factors. Therefore, further investigation into these potential solutions is warranted. 


The following are some helpful research studies. The contents of these studies relate to the contents briefly mentioned above:

- [Study of the Effects of Alcohol on Drivers and Driving Performance on Straight Road](https://www.hindawi.com/journals/mpe/2014/607652/)
- [Preventative Legislation for Drinking and Driving](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6876524/)



# Background
Data was collected for 48 states in the United States, with Alaska and Hawaii being the two excluded states. The traffic fatality data was gathered from the US Department of Transportation Fatal Accident Reporting System (“Fatality Analysis Reporting System (FARS)” 2016) which is a census conducted nationwide to provide the public with data related to the traffic crashes. The data set contains 336 entries and 34 variables. The number of entries is 336, since there is an entry for each of the 48 states for the years 1982 through 1988 (“Fatalities” n.d.). 

- Explanation of variables found here: [Fatalities](https://www.rdocumentation.org/packages/AER/versions/1.2-9/topics/Fatalities)
- Info about the FARS: [Fatality Analysis Reporting System](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars)

Given the goal of the analysis is to help determine policy and legislative changes that would be of interest to legislators at the state and federal level. Some policies of interest present in the dataset include the legal drinking age (ranging from 18-21), presence of a preliminary breath test law, as well as mandatory jail sentence for alcohol induced traffic accidents. Other data available for potential exploration are spirits consumption score for regions, which is a measure of high percentage alcohol content intake (>30% alcohol by volume (ABV). In addition, the tax on a case of beer is provided to see if taxation rates can influence fatal traffic incidents.

There is interest in further exploring variables related to the amount of travel in a region, measured in average miles per driver. In addition, other socio-economic factors like the amount of young drivers and income with regards to traffic incidents will be explored. Some of the response variables of interest are the fatality rate which can be obtained using the total vehicle fatalities after normalizing by population size. Similar methods can be used to explore the overall rates for the age groups 15-17, 18-20, and 21-24.



# Methods/Results

## Descriptive Analysis


---
title: "Project 2"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r echo=FALSE, warning=FALSE,  results='asis', include=FALSE}
library(AER)
data('Fatalities')
```
```{r echo=FALSE, warning=FALSE,  results='asis', include=FALSE}
# - Explanation of variables found here: [Fatalities](https://www.rdocumentation.org/packages/AER/versions/1.2-9/topics/Fatalities)


library(tidyverse)
library(lubridate)
library(timetk)
library(kableExtra)
library(devtools)
library(gtsummary)
library(plotly)
library(DT)
library(tidyverse)
library(lubridate)
```


```{r echo=FALSE, warning=FALSE,  results='asis'}
Fatalities$frate = Fatalities$fatal/Fatalities$pop * 10000
ggplot(data = Fatalities, aes(x = year, y = fatal, group = state, colour = state)) +
    geom_line()
```

```{r echo=FALSE, warning=FALSE,  results='asis', include=FALSE}
data1 = Fatalities[1:112, ]
data2 = Fatalities[113:224, ]
data3 = Fatalities[225:336, ]
ggplot(data = data1, aes(x = year, y = fatal, group = state, colour = state)) +
    geom_line() +
    facet_wrap(~ state) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90), panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank())
ggplot(data = data2, aes(x = year, y = fatal, group = state, colour = state)) +
    geom_line() +
    facet_wrap(~ state) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90), panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank())

ggplot(data = data3, aes(x = year, y = fatal, group = state, colour = state)) +
    geom_line() +
    facet_wrap(~ state) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90), panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank())
#We first make the time series plots for the number of fatalities of each state from 1982 to 1988. From the plots above, we can find that the numbers of fatalities vary a lot, making it difficult to identify if there is any pattern. Therefore, we try to use the ratio of the number of fatalities over the population of the state, and make the time series plots for each state again. From the plots below, we can observe that there are certain states show similar patterns. Therefore, we want to understand what the factors in this data set affect the change of the number of fatalities are. 
```



```{r echo=FALSE, warning=FALSE,  results='asis'}
##Time series plots using the ratio of the total number of fatalities over the population of the state
#Fatalities$frate = Fatalities$fatal/Fatalities$pop
data4 = Fatalities[1:112, ]
data5 = Fatalities[113:224, ]
data6 = Fatalities[225:336, ]
ggplot(data = data4, aes(x = year, y = frate, group = state, colour = state)) +
    geom_line() +
    facet_wrap(~ state) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90), panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank())
ggplot(data = data5, aes(x = year, y = frate, group = state, colour = state)) +
    geom_line() +
    facet_wrap(~ state) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90), panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank())

ggplot(data = data6, aes(x = year, y = frate, group = state, colour = state)) +
    geom_line() +
    facet_wrap(~ state) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90), panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank())
```





```{r echo=FALSE, warning=FALSE,  results='asis', include=FALSE}
par(mfrow=c(1,2))
jail_yes = subset(Fatalities, jail == 'yes') 
jail_no = subset(Fatalities, jail == 'no') 
hist(jail_no$fatal, main = 'Fatalities without mandatory jail', , xlab = "No Jail")
hist(jail_yes$fatal,  main = 'Fatalities with mandatory jail', xlab = "Jail")
#mean(jail_no$fatal)
#mean(jail_yes$fatal)
```

We are interested to know  if the mandatory jail sentence for the traffic fatalities has any relationship with the number of fatalities. From the histogram of the number of fatalities with mandatory jail sentence and that without, we find that the the number of fatalities with mandatory jail sentence is smaller than that without jail sentence. 


```{r echo=FALSE, warning=FALSE,  results='asis'}
par(mfrow=c(1,2))
jail_yes = subset(Fatalities, jail == 'yes') 
jail_no = subset(Fatalities, jail == 'no') 
hist(jail_no$frate, main = 'Fatalities without mandatory jail', , xlab = "No Jail")
hist(jail_yes$frate,  main = 'Fatalities with mandatory jail', xlab = "Jail")
#mean(jail_no$fatal)
#mean(jail_yes$fatal)
```

However, because the populations of different states vary, we also use the ratio of the total number of fatalities over the population of the state to plot histograms again, and we find that the ratio in the histogram of mandatory jail sentence is larger than that that in the histogram without jail sentence. So, the population of each state may be affecting our analysis. 


```{r echo=FALSE, warning=FALSE,  results='asis'}
Fatalities$drinkage = round(Fatalities$drinkage)
Fatalities$drinkage = as.factor(Fatalities$drinkage)
```

```{r echo=FALSE, warning=FALSE,  results='asis', include=FALSE}
ggplot(Fatalities, aes(x=drinkage, y=fatal)) + 
  geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0))

#We also investigate the relationship between the number of fatalities and the minimum legal drinking age. Since the minimum drinking ages are not necessarily whole numbers, we round them to the nearest integer. From the twp-way box plot above, we can observe that the states with legal minimum drinking ages 18 and 21 are associated with the largest average number of fatalities, and that the states with minimum legal drinking age 20 seems has the smallest average number of fatalities. We also noticed that the data points of minimum drinking age 21 disperse widely.   

```


```{r echo=FALSE, warning=FALSE,  results='asis', include=FALSE}
test <- Fatalities %>% select(unemp, frate, jail)
test %>% tbl_summary(by=jail, digits = all_continuous() ~ 6, statistic = list(frate ~ "{mean} ({sd}, {median}, {IQR})")) %>% add_p()


test <- Fatalities %>% select(frate, drinkage)
test %>% tbl_summary(by=drinkage, digits = all_continuous() ~ 6, statistic = list(frate ~ "{mean} ({sd}, {median}, {IQR})")) %>% add_p()
```

```{r echo=FALSE, warning=FALSE,  results='asis', include=FALSE}
plot(Fatalities$milestot, Fatalities$frate, xlab = 'Total vechicle miles', ylab = 'The fatality rate' )
#From the plot above, we can find that the there is a linear relationship between the number of fatalities and total vehicle miles. 

```



```{r echo=FALSE, warning=FALSE,  results='asis'}
par(mfrow=c(1,2))
plot(Fatalities$income, Fatalities$fatal, xlab = 'Income', ylab = 'Number of fatalities')
plot(Fatalities$income, Fatalities$frate, xlab = 'Income', ylab = 'Fatality Rate')
```

From the plots above, we can not find any relationship between the number of fatalities and income. However, if we use the ratio of the total number of fatalities over the population of the state, we can find there is a quadratic pattern between income and the ratio.


```{r echo=FALSE, warning=FALSE,  results='asis'}
Fatalities$inc_cat = ifelse(Fatalities$income <= 12086, '1st',ifelse(Fatalities$income > 12086 & Fatalities$income <=13763, '2nd',ifelse(Fatalities$income >13763 & Fatalities$income <=15175, '3rd', '4th')))
```

```{r echo=FALSE, warning=FALSE,  results='asis'}
#test <- Fatalities %>% select(frate, bina)
#test %>% tbl_summary(by=inc_cat, digits = all_continuous() ~ 6, statistic = list(frate ~ "{mean} ({sd}, {median}, {IQR})")) %>% add_p()
```

```{r echo=FALSE, warning=FALSE,  results='asis'}
boxplot(Fatalities$frate~Fatalities$inc_cat, xlab = 'Quantile of income', ylab = 'Fatality Rate')
```

We use quantiles to reclassify income, and plot the two-way box plot between income and the fatality rate. We can observe that the third and fourth quantiles of incomes appear to havehave lower average fatality rate.
```{r echo=FALSE, warning=FALSE,  results='asis'}
#summary(Fatalities$unemp)

test <- Fatalities %>% select(inc_cat, frate)
test %>% tbl_summary(by=inc_cat, digits = all_continuous() ~ 6, statistic = list(frate ~ "{mean} ({sd}, {median}, {IQR})")) %>% add_p()
```

```{r echo=FALSE, warning=FALSE,  results='asis'}
Fatalities$unem_cat = ifelse(Fatalities$unemp <= 5.475, '1st',ifelse(Fatalities$unemp > 5.475  & Fatalities$unemp <=7.000, '2nd',ifelse(Fatalities$unemp >7.000 & Fatalities$unemp <=8.900, '3rd', '4th')))
```

```{r echo=FALSE, warning=FALSE,  results='asis'}
boxplot(Fatalities$frate~Fatalities$unem_cat, xlab = 'Quantile of unemployment rate', ylab = 'Fatality Rate')
```

We use quantiles to reclassify unemployment rate, and plot two-way box plot between unemployment rate and the fatality rate, we can observe that the third and the first quantiles have the lowest average numbers of fatalities.

```{r echo=FALSE, warning=FALSE,  results='asis'}
#summary(Fatalities$unemp)

test <- Fatalities %>% select(unem_cat, frate)
test %>% tbl_summary(by=unem_cat, digits = all_continuous() ~ 6, statistic = list(frate ~ "{mean} ({sd}, {median}, {IQR})")) %>% add_p()
```


```{r echo=FALSE, warning=FALSE,  results='asis'}
#write.csv(Fatalities, 'post_EDA.csv')
```




```{r loadLibs,echo=FALSE,warning=FALSE,message=FALSE}
library(AER)
library(tidyverse)
library(lubridate)
library(lmerTest)
library(scales)
library(kableExtra)
library(ggpubr)
```
```{r loadData, echo=FALSE}
data('Fatalities')
```
```{r briefView, echo=FALSE}
summary(Fatalities[, c('year', 'state', 'pop', 'fatal', 'income', 'unemp', 'miles', 'beertax', 'youngdrivers',  'jail', 'breath', 'service')])
Fatalities[which(is.na(Fatalities$jail) | is.na(Fatalities$service)), c('state','year', 'jail', 'service')]
```
Notice California has missing values in `jail` and `service` only in year 1988.
We create a new variable "frate" which is the the number of fatalities per 10000 residents. 

```{r addVar, echo = FALSE}
Fatalities$frate = Fatalities$fatal/Fatalities$pop * 10000
```

```{r jail_vs_covariate, echo=FALSE, warning=FALSE}

# explore  relationships between jail and unemployment
p1 = ggplot(data=Fatalities, aes(unemp, fill = jail)) + 
          geom_density( alpha = 0.5 ) + 
          labs(title = "jail vs unemployment")
# explore relationships between jail and income
p2 = ggplot(data=Fatalities, aes(income, fill = jail)) + 
          geom_density( alpha = 0.5 )+ 
  labs(title = "jail vs income")


# explore relationships between jail and miles
p3 = ggplot(data=Fatalities, aes(miles, fill = jail)) + 
          geom_density( alpha = 0.5 )+
  labs(title = "jail vs miles")

# explore relationships between jail and beertax
p4 =ggplot(data=Fatalities, aes(beertax, fill = jail)) + 
          geom_density( alpha = 0.5 ) + 
  labs(title = "jail vs beertax")
# explore relationships between jail and youngdrivers
p5 =ggplot(data=Fatalities, aes(youngdrivers, fill = jail)) + 
          geom_density( alpha = 0.5 ) +
   labs(title = "jail vs youngdrivers")


ggarrange(p1, p2, p3, p4, p5, labels = LETTERS[1:5],
          common.legend = TRUE, legend = "bottom")

```

- unemployment does not seem to strongly impact the propensity of being assigned to treatment (jail = yes) or control (jail = no) group.

- It seems that income impacts the propensity of being assigned to treatment (jail = yes) or control (jail = no) group. Therefore income should be included in the selection model.

- mileage driven per person does not seem to strongly impact the propensity of being assigned to treatment (jail = yes) or control (jail = no) group.

- beertax does not seem to strongly impact the propensity of being assigned to treatment (jail = yes) or control (jail = no) group.

- percent of young drivers does not seem to strongly impact the propensity of being assigned to treatment (jail = yes) or control (jail = no) group.

```{r coolFun, echo=FALSE, warning=FALSE}
# Adpated from http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
my_cols <- c("#00AFBB", "#FC4E07")  

# Correlation panel
panel.cor <- function(x, y){
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y)
    txt <- paste0("r = ", round(r, digits = 2))
    text(0.5, 0.5, txt, cex = 3*abs(r) )
}

# Customize upper panel
upper.panel<-function(x, y){
  points(x,y, pch = 19, col = alpha( my_cols[Fatalities$jail], 0.3 ) )
}

```

```{r echo=FALSE, warning=FALSE}
pairs( Fatalities[, c('frate', 'income', 'unemp', 'miles', 'beertax', 'youngdrivers') ],
       upper.panel = upper.panel,
       lower.panel = panel.cor )
```


Independent variables chosen based on the degree of correlation with "frate", and we start by building a tentative outcome model without the use of propensity scores.



## Predictive/Inferential Analysis


---

The package lme4 for Fitting Linear Mixed-Effects Models was utilized to allow for both fixed and random effects. Due to multiple observations for each state from the years 1982-1988 there is inherent non-randomness in our data. Therefore, this analysis is interested in variability for each state and its relative fatality rate and this is preferred to aggregation of the data as it does not properly utilize all of the data. 

- [lme4 Package Documentation](https://cran.r-project.org/web/packages/lme4/vignettes/lmer.pdf)


The Linear Mixed Model follows the following form:
$$y = \textbf{X}\beta + Z\mu + \epsilon$$

- $y$ and $\epsilon$ are both nx1 vectors representing the dependent variable and the residuals. The residuals refer to the amount of the dependent variable, in this case fatality rate, that is not explained by the model. 
- $n$ is the number of observations in the dataset, in this case 336. 
- $\textbf{X}$ is an $n$ x $p$ matrix where there are $p$ predictor variables in the model. 
- $\textbf{Z}$ is the random effects design matrix of size $N$ x $qJ$ where there are $q$ random effects of interest and $J$ groups. (Vasishth and Broe 2011)

---

### Model 1 (treatment (jail = yes) and control (jail = no))

```{r passModel1, echo=FALSE, warning=FALSE}

summary( lmer(frate ~ jail + rescale(unemp) + rescale(income) + rescale(beertax) + rescale(youngdrivers) +  rescale(miles) + (1|state), data=Fatalities) )$coef %>% round(.,3) %>%  kable() %>% kable_styling()

```

We found a non-significant treatment effect, as the factor level jailyes is not statistically significant at p = 0.05. However, the covariates `unemp` and `youngdrivers` are statistically significant at p < 0.001.

Based on our exploratory data analysis and group discussion, we assume that the variables `unemp`, `income`, `beertax`, `youngdrivers` and `miles` have an influence on the assignment to either the control (jail = no) or the treatment (jail = yes) group.

Next we performed a balance analysis, trying to assess the degree of bias during assignment to treatment and control groups, before the use of propensity scores.


```{r preBalance1, echo = FALSE, warning=FALSE, results='asis'}
cat("#### Balance analysis before estimating propensity scores:\n")
cat("- jail vs. miles\n")

summary( lm(miles ~ jail, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. income\n")
summary( lm(income ~ jail, data = Fatalities) ) $coefficients %>% round(., 5)%>% kable() %>% kable_styling()

cat("- jail vs. beertax\n")
summary( lm(beertax ~ jail, data = Fatalities) )$coefficients %>% round(., 5)%>% kable() %>% kable_styling()

cat("- jail vs. youngdrivers\n")
summary( lm(youngdrivers ~ jail, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. unemployment\n")
summary( lm(unemp ~ jail, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()
```
This time it turns out that covariate `unemp` shows strong imbalance between treatment (jail = yes) and control (jail = no ) groups (p < 0.01), and covariate `income` shows a marginal significant imbalance (p < 0.05). Although not significant, the significance levels of covariates `miles` and `beertax` are close to p = 0.05 level, also suggesting potential selection bias. These should be addressed using propensity scores.

```{r selMod1, echo=FALSE, warning=FALSE}
mod1 = glm(jail ~ unemp + income + beertax + youngdrivers + miles,
           family="binomial", data=Fatalities)
summary(mod1)$coef %>% round(., 5) %>% kable() %>% kable_styling()
```
As expected from descriptive analysis, we found that covariate `income` significantly impact the assignment to treatment or control group at p < 0.05. Note that the significance level of `unemp` does not reach p = 0.1, probably due to its strong negative correlation with `income` (r = -0.55) and therefore does not contribute more information when `income` is included in the model. We decide to keep all 5 variables in the selection model regardless of their significance level, as advocated by Caliendo and Kopeinig (2008).




```{r outcomeMod1, echo=FALSE, warning=FALSE}
prob1 = mod1$fitted.values
pscore1 = ifelse( Fatalities$jail == "yes", prob1,(1-prob1) )
weight1 = 1/pscore1

mod1_new = lmer(frate ~ jail + rescale(unemp) + rescale(income) + rescale(beertax) + rescale(youngdrivers) + rescale(miles) + (1|state), data=Fatalities, weights=weight1)

summary(mod1_new)$coef %>% round(., 5) %>% kable() %>% kable_styling()
```
We used logistic regression to determine the probability of treatment/control assignment, given the 5 chose covariates mentioned earlier. The estimated probablities were then converted to propensity scores, whose inverse are included as weights in the linear mixed model.
After applying propensity score weighting, we still do not found significant treatment effect, as the factor level jailyes is not statistically significant at p = 0.05. Again, the covariates `unemp` and `youngdrivers` are statistically significant at p < 0.001.

```{r postBal1, echo=FALSE, warning=FALSE, results='asis'}
cat("#### Balance analysis after estimating propensity scores:\n")
cat("- jail vs. miles\n")
summary( lm(miles ~ jail, weight = weight1, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. income\n")
summary( lm(income ~ jail, weight = weight1,data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. beertax\n")
summary( lm(beertax ~ jail, weight = weight1,data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. youngdrivers\n")
summary( lm(youngdrivers ~ jail, weight = weight1, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. unemployment\n")
summary( lm(unemp ~ jail, weight = weight1, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()
```

The selection biases by `income` and `unemp` are reduced much after propensity score weighting.


### Model 2 (treatment (breath = yes) and control (breath = no))

```{r passModel2, echo = FALSE,warning=FALSE}

summary( lmer(frate ~ breath + rescale(unemp) + rescale(income) + rescale(beertax) + rescale(youngdrivers) +  rescale(miles) + (1|state), data=Fatalities) )$coef %>% round(.,3) %>%  kable() %>% kable_styling()

```

We found a marginally treatment effect, as the factor level breathyes is statistically significant at p = 0.05. Interestingly, the covariates `unemp` and `youngdrivers` are statistically significant at p < 0.001.

Likewise, we performed a balance analysis before the use of propensity scores.



```{r preBal2, echo = FALSE,warning=FALSE, results='asis'}
cat("#### Balance analysis before estimating propensity scores:\n")
cat("- jail vs. miles\n")
summary( lm(miles ~ breath, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. income\n")
summary( lm(income ~ breath, data = Fatalities) ) $coefficients %>% round(., 5)%>% kable() %>% kable_styling()

cat("- jail vs. beertax\n")
summary( lm(beertax ~ breath, data = Fatalities) )$coefficients %>% round(., 5)%>% kable() %>% kable_styling()

cat("- jail vs. youngdrivers\n")
summary( lm(youngdrivers ~ breath, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. unemployment\n")
summary( lm(unemp ~ breath, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()
```

This time it turns out that covariate `unemp` shows strong imbalance between treatment (jail = yes) and control (jail = no ) groups (p < 0.01), and covariate `income` shows a marginal significant imbalance (p < 0.05). Although not significant, the significance levels of covariates `miles` and `beertax` are close to p = 0.05 level, also suggesting potential selection bias. These should be addressed using propensity scores.

```{r selModel2, echo = FALSE,warning=FALSE}
mod2 = glm(breath ~ unemp + income + beertax + youngdrivers + miles,
           family="binomial", data=Fatalities)
summary(mod2)$coef %>% round(., 5) %>% kable() %>% kable_styling()
```
 We found that covariate `unemp` significantly impact the assignment to treatment or control group at p < 0.001, and covariate `miles` significantly impact the assignment to treatment or control group at p < 0.01. Again, all 5 variables are included in the selection model regardless of their significance level.
 
```{r outcomeModel2, echo = FALSE,warning=FALSE}
prob2 = mod2$fitted.values
pscore2 = ifelse(Fatalities$breath == "yes", prob2,(1-prob2))
weight2 = 1/pscore2
mod2_new = lmer( frate ~ breath+rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+rescale(miles)+(1|state),data=Fatalities, weights=weight2)
summary(mod2_new)$coef %>% round(., 5) %>% kable() %>% kable_styling()
```
Likewise, after applying propensity score weighting, we still do not found significant treatment effect, as the factor level breathyes is not statistically significant at p = 0.05. Nevertheless, it is significant at p < 0.1, indicating potential effects that needs further investigation. Again, the covariates unemp and youngdrivers are statistically significant at p < 0.001.

```{r postBal2, echo = FALSE,warning=FALSE, results='asis'}
cat("#### Balance analysis after estimating propensity scores:\n")
cat("- jail vs. miles\n")
summary( lm(miles ~ breath, weight = weight2, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. income\n")
summary( lm(income ~ breath, weight = weight2,data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. beertax\n")
summary( lm(beertax ~ breath, weight = weight2,data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. youngdrivers\n")
summary( lm(youngdrivers ~ breath, weight = weight2, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. unemployment\n")
summary( lm(unemp ~ breath, weight = weight2, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()
```
Note that the selection biases by `income` and `unemp` are not are reduced much after propensity score weighting, so there still exists substantial selection biases that may invalidate causal inferences.

### Model 3 (treatment (service = yes) and control (service = no))

```{r selModel3, echo = FALSE,warning=FALSE }
mod3 = glm(service ~ rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+rescale(miles),
           family="binomial", data=Fatalities)
prob3 = mod3$fitted.values
pscore3 = ifelse(Fatalities$service == "yes", prob3,(1-prob3))
weight3 = 1/pscore3
mod3_new = lmer(frate~service+rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+
                  rescale(miles)+(1|state),data=Fatalities, weights=weight3)
summary(mod3_new)$coefficients %>% round(., 5) %>% kable() %>% kable_styling()
```

Using the same method, we do not found significant treatment effect, as the factor level serviceyes is not statistically significant at p = 0.05 Again, the covariates `unemp` and `youngdrivers` are statistically significant at p < 0.001.

```{r balance3, echo = FALSE,warning=FALSE, results='asis' }
cat("#### Balance analysis before estimating propensity scores:\n")

cat("- jail vs. miles\n")
summary( lm(miles ~ service, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. income\n")
summary( lm(income ~ service,data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. beertax\n")
summary( lm(beertax ~ service, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. youngdrivers\n")
summary( lm(youngdrivers ~ service, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. unemployment\n")
summary( lm(unemp ~ service, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("#### Balance analysis after estimating propensity scores:\n")
cat("- jail vs. miles\n")
summary( lm(miles ~ service, weight = weight3, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. income\n")
summary( lm(income ~ service, weight = weight3,data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. beertax\n")
summary( lm(beertax ~ service, weight = weight3,data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. youngdrivers\n")
summary( lm(youngdrivers ~ service, weight = weight3, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()

cat("- jail vs. unemployment\n")
summary( lm(unemp ~ service, weight = weight3, data = Fatalities) )$coefficients %>% round(., 5) %>% kable() %>% kable_styling()
```

It seems that the selection biases by `beertax` are effectively removed after propensity score weighting.



# Acknowledgements

Dr. Shizhe Chen and Yejiong Zhu for their course materials and instruction for data analysis techniques and R code. 




# Supplementary Material

Github link for code is provided to reproduce visualizations and statistical analysis from the study, in addition an HTML is provided with pre-computed results: https://github.com/keithgmitchell/STA207GroupProjects





# References

Caliendo, Marco, and Sabine Kopeinig. 2008. “Some Practical Guidance for the Implementation of Propensity Score Matching.” Journal of Economic Surveys 22 (1): 31–72.

“Fatalities.” n.d. Accessed February 11, 2021. https://www.rdocumentation.org/packages/AER/versions/1.2-9/topics/Fatalities.

“Fatality Analysis Reporting System (FARS).” 2016. November 14, 2016.
https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars.

Hingson, Ralph. 1996. “Prevention of Drinking and Driving.” Alcohol Health and Research World 20 (4): 219–26.

Sehat, Mojtaba, Kourosh Holakouie Naieni, Mohsen Asadi-Lari, Abbas Rahimi Foroushani, and Hossein Malek-Afzali. 2012. “Socioeconomic Status and Incidence of Traffic Accidents in Metropolitan Tehran: A Population-Based Study.” International Journal of Preventive Medicine 3 (3): 181–90.

Vasishth, Shravan, and Michael Broe. 2011. “An Introduction to Linear Mixed Models.” In The Foundations of Statistics: A Simulation-Based Approach, 145–59. Berlin, Heidelberg: Springer Berlin Heidelberg.

Zhao, Xiaohua, Xingjian Zhang, and Jian Rong. 2014. “Study of the Effects of Alcohol on Drivers and Driving Performance on Straight Road.” Mathematical Problems in Engineering 2014 (February). https://doi.org/10.1155/2014/607652.
