---
title: "Project 2 Post Exploration"
author: "Keith Mitchell"
date: "2/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MASS)
library(plotrix)
library(tidyverse)
library(lubridate)
library(timetk)
library(kableExtra)
library(devtools)
library(gtsummary)
library(plotly)
library(DT)
```

## TODO manage outliers and test with training and non training set. (other code from 206 for comparing STEP aic number of variable to be chosen max)

```{r}
fatalities = read_csv('post_EDA.csv')
```
## Lets explore predictors (that make sense) and are interpretable for overall fatalities
- We will call this the general model
```{r}
drops <- c("nfatal","sfatal", "afatal", "state", "X1","frate")
fatalities_gen = fatalities[,!grepl("1517", colnames(fatalities))]
fatalities_gen = fatalities_gen[,!grepl("1820", colnames(fatalities_gen))]
fatalities_gen = fatalities_gen[,!grepl("2124", colnames(fatalities_gen))]
fatalities_gen = fatalities_gen[,!grepl("2124", colnames(fatalities_gen))]
fatalities_gen = fatalities_gen[, !(names(fatalities_gen) %in% drops)]


colnames(fatalities_gen)
fatalities_gen
fatalities_gen <- na.omit(fatalities_gen)
fatalities_gen
```

```{r}
null = lm(fatal ~ 1, data=fatalities_gen)
full = lm(fatal ~ ., data=fatalities_gen)
step.f<-stepAIC(null, scope = list(lower=null,upper=full), direction = 'forward', trace=FALSE)

summary(step.f)
datatable(confint.default(step.f)[-c(1),])
par(mfrow=c(2,2))
plot(step.f)
```
```{r}
ci = confint.default(step.f)

df <- data.frame(x =rownames(ci[-c(1),]),
                 F =step.f$coefficients[-c(1)],
                 L =confint.default(step.f)[-c(1),][,1],
                 U =confint.default(step.f)[-c(1),][,2])
ggplot(df, aes(x = x, y = F)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymax = U, ymin = L))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


## Lets explore predictors (that make sense) and are interpretable for alcohol fatalities
- We will call this the alcohol fatalities model

```{r}
drops <- c("nfatal","sfatal", "fatal", "state", "X1", "frate")
fatalities_a = fatalities[,!grepl("1517", colnames(fatalities))]
fatalities_a = fatalities_a[,!grepl("1820", colnames(fatalities_a))]
fatalities_a = fatalities_a[,!grepl("2124", colnames(fatalities_a))]
fatalities_a = fatalities_a[,!grepl("2124", colnames(fatalities_a))]
fatalities_a = fatalities_a[, !(names(fatalities_a) %in% drops)]

colnames(fatalities_a)
fatalities_a
fatalities_a <- na.omit(fatalities_a)
fatalities_a
```

```{r}
null = lm(afatal ~ 1, data=fatalities_a)
full = lm(afatal ~ ., data=fatalities_a)
step.f<-stepAIC(null, scope = list(lower=null,upper=full), direction = 'forward', trace=FALSE)
summary(step.f)
datatable(confint.default(step.f))
par(mfrow=c(2,2))
plot(step.f)

```


```{r}
ci = confint.default(step.f)

df <- data.frame(x =rownames(ci[-c(1),]),
                 F =step.f$coefficients[-c(1)],
                 L =confint.default(step.f)[-c(1),][,1],
                 U =confint.default(step.f)[-c(1),][,2])
ggplot(df, aes(x = x, y = F)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymax = U, ymin = L))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```



## Lets explore predictors (that make sense) and are interpretable for frate
- We will call this the alcohol fatalities model
- TODO should we do frate for the afatal only??
- TODO outliers need to be managed.
```{r}
drops <- c("nfatal","sfatal", "afatal", "X1","fatal","state")
fatalities_rate = fatalities[,!grepl("1517", colnames(fatalities))]
fatalities_rate = fatalities_rate[,!grepl("1820", colnames(fatalities_rate))]
fatalities_rate = fatalities_rate[,!grepl("2124", colnames(fatalities_rate))]
fatalities_rate = fatalities_rate[,!grepl("2124", colnames(fatalities_rate))]
fatalities_rate = fatalities_rate[, !(names(fatalities_rate) %in% drops)]


colnames(fatalities_rate)
fatalities_rate
fatalities_rate <- na.omit(fatalities_rate)
fatalities_rate
boxplot(fatalities_rate$spirits)

```

```{r}
null = lm(frate ~ 1, data=fatalities_rate)
full = lm(frate ~ ., data=fatalities_rate)
summary(full)
```

```{r}
step.f<-stepAIC(null, scope = list(lower=null,upper=full), direction = 'forward', trace=FALSE)
summary(step.f)
```
```{r}
anova(step.f)
datatable(confint.default(step.f)[-c(1),])
par(mfrow=c(2,2))
plot(step.f)
```
```{r}
fatalities_rate[170,]
data("Produc", package = "plm")
zz <- plm(log(gsp) ~ log(pcap) + log(pc) + log(emp) + unemp,
          data = Produc, index = c("state","year"))
summary(zz)
```
```{r}
ci = confint.default(step.f)

df <- data.frame(x =rownames(ci[-c(1),]),
                 F =step.f$coefficients[-c(1)],
                 L =confint.default(step.f)[-c(1),][,1],
                 U =confint.default(step.f)[-c(1),][,2])
ggplot(df, aes(x = x, y = F)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymax = U, ymin = L))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

```{r}
Fatalities
```

```{r}
mod1 = glm(jail ~ rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+rescale(miles), family="binomial", data=Fatalities)
prob1 = mod1$fitted.values
pscore1 = ifelse(Fatalities$jail == "yes", prob1,(1-prob1))
library("devtools"); install_github("lme4/lme4",dependencies=TRUE)

mod1_new = lmer(frate~jail+rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+rescale(miles)+(1|state),data=Fatalities, weights=weight)
summary(mod1_new)
```