---
title: "Project 2"
author: "Seyoung Jung"
date: "2/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r,warning=FALSE,message=FALSE}
library(AER)
library(tidyverse)
library(lubridate)
library(lmerTest)
library(scales)
library(kableExtra)
```


```{r}
data('Fatalities')
```


```{r}
summary(Fatalities)
```

Observe the row(s) containing NA value in "jail" and/or "service".
One row (row #28) has NA in the two variables.
```{r}
Fatalities[which(is.na(Fatalities$jail) | is.na(Fatalities$service)), ]
```


Observe the row which has the maximum value from the "miles" variable
Notice the stark difference between the highest and the second highest average miles per driver 
```{r}
Fatalities[order(-Fatalities$miles), ] %>% head(5)

plot(Fatalities$miles[-171], Fatalities$fatal[-171], xlab = 'Average miles per driver', ylab = 'The number of fatalities' )

plot(Fatalities$miles[-171], Fatalities$frate[-171], xlab = 'Average miles per driver', ylab = 'Fatality rate' )
```


Check whether the other rows of California have data on "jail" and "service".
The other rows have "no"s in the jail and service columns. 
```{r}
Fatalities[which(Fatalities$state=="ca"), ]
```


Convert the "state", "year", "drinkage" into a factor variable
```{r}
Fatalities$drinkage = round(Fatalities$drinkage)
Fatalities$drinkage = as.factor(Fatalities$drinkage)
```


Correlation between "fatal" and the other non-categorical variables
```{r}
cor(Fatalities$fatal, Fatalities[, -which(sapply(Fatalities, class)=="factor")])
```


Create a new variable "frate" which is the total number of fatalities over the population of the state. 
Correlation between "frate" and the other non-categorical variables
Some variables (e.g. income, beertax, baptist, miles, and youngdrivers) have relatively high correlations with "frate".
```{r}
Fatalities$frate = Fatalities$fatal/Fatalities$pop

cor(Fatalities$frate, Fatalities[, -which(sapply(Fatalities, class)=="factor")])
```








```{r}
#summary(glm(breath~frate,data=Fatalities,family=binomial))$coefficients # breath vs. frate
#summary(glm(jail~frate,data=Fatalities,family=binomial))$coefficients # jail vs. frate
#summary(glm(service~frate,data=Fatalities,family=binomial))$coefficients # service vs. frate

#summary(glm(breath~unemp,data=Fatalities,family=binomial))$coefficients # breath vs. unemp
#summary(glm(jail~unemp,data=Fatalities,family=binomial))$coefficients # jail vs. unemp
#summary(glm(service~unemp,data=Fatalities,family=binomial))$coefficients # service vs. unemp

#summary(glm(breath~miles,data=Fatalities,family=binomial))$coefficients # breath vs. miles
#summary(glm(jail~miles,data=Fatalities,family=binomial))$coefficients # jail vs. miles
#summary(glm(service~miles,data=Fatalities,family=binomial))$coefficients # service vs. miles

#summary(glm(breath~youngdrivers,data=Fatalities,family=binomial))$coefficients # breath vs. youngdrivers
#summary(glm(jail~youngdrivers,data=Fatalities,family=binomial))$coefficients # jail vs. youngdrivers
#summary(glm(service~youngdrivers,data=Fatalities,family=binomial))$coefficients # service vs. youngdrivers

#summary(glm(breath~beertax,data=Fatalities,family=binomial))$coefficients # breath vs. beertax
#summary(glm(jail~beertax,data=Fatalities,family=binomial))$coefficients # jail vs. beertax
#summary(glm(service~beertax,data=Fatalities,family=binomial))$coefficients # service vs. beertax

#summary(glm(breath~income,data=Fatalities,family=binomial))$coefficients # breath vs. income
#summary(glm(jail~income,data=Fatalities,family=binomial))$coefficients # jail vs. income
#summary(glm(service~income,data=Fatalities,family=binomial))$coefficients # service vs. income
```


```{r}
# explore  relationships between jail and unemployment
ggplot(data=Fatalities, aes(unemp, fill = jail)) + 
          geom_density( alpha = 0.5 )
```
unemployment does not seem to strongly impact the propensity of being assigned to treatment (jail = yes) or control (jail = no) group.

```{r}
# explore relationships between jail and income
ggplot(data=Fatalities, aes(income, fill = jail)) + 
          geom_density( alpha = 0.5 )

```
It seems that income impacts the propensity of being assigned to treatment (jail = yes) or control (jail = no) group. Therefore income should be included in the selection model.

```{r}
# explore relationships between jail and fatalities
ggplot(data=Fatalities, aes(miles, fill = jail)) + 
          geom_density( alpha = 0.5 )

```
mileage driven per person does not seem to strongly impact the propensity of being assigned to treatment (jail = yes) or control (jail = no) group.

```{r}
# explore relationships between jail and beertax
ggplot(data=Fatalities, aes(beertax, fill = jail)) + 
          geom_density( alpha = 0.5 )
```

beertax does not seem to strongly impact the propensity of being assigned to treatment (jail = yes) or control (jail = no) group.

```{r}
# explore relationships between jail and youngdrivers
ggplot(data=Fatalities, aes(youngdrivers, fill = jail)) + 
          geom_density( alpha = 0.5 )
```

percent of young drivers does not seem to strongly impact the propensity of being assigned to treatment (jail = yes) or control (jail = no) group.

```{r}
# Adpated from http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
my_cols <- c("#00AFBB", "#FC4E07")  

# Correlation panel
panel.cor <- function(x, y){
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y)
    txt <- paste0("r = ", round(r, digits = 2))
    text(0.5, 0.5, txt, cex = 3*abs(r) )
}

# Customize upper panel
upper.panel<-function(x, y){
  points(x,y, pch = 19, col = alpha( my_cols[Fatalities$jail], 0.3 ) )
}

```

```{r}
pairs( Fatalities[, c('income', 'unemp', 'miles', 'beertax', 'youngdrivers') ],
       upper.panel = upper.panel,
       lower.panel = panel.cor )
```



==========================================================================================
                              Model Building (Seyoung Jung)
==========================================================================================

```{r}
 cor(Fatalities$frate, Fatalities[, -which(sapply(Fatalities, class)=="factor")])  


```
Independent variables chosen based on the degree of correlation with "frate"

We start by building a tentative outcome model without the use of propensity scores.

## Model 1 (treatment (jail = yes) and control (jail = no))

```{r warning=FALSE}

summary( lmer(frate ~ jail + rescale(unemp) + rescale(income) + rescale(beertax) + rescale(youngdrivers) +  rescale(miles) + (1|state), data=Fatalities) )$coef %>% round(.,3) %>%  kable() %>% kable_styling()

```

We found a non-significant treatment effect, as the factor level jailyes is not statistically significant at p = 0.05. However, the covariates `unemp` and `youngdrivers` are statistically significant at p < 0.001.

Based on our exploratory data analysis and group discussion, we assume that the variables `unemp`, `income`, `beertax`, `youngdrivers` and `miles` have an influence on the assignment to either the control (jail = no) or the treatment (jail = yes) group.

Next we performed a balance analysis, trying to assess the degree of bias during assignment to treatment and control groups, before the use of propensity scores.

```{r}
# balance analysis before estimating propensity scores
summary( lm(miles ~ jail, data = Fatalities) )$coefficients %>% round(., 2) %>% kable() %>% kable_styling()
summary( lm(income ~ jail, data = Fatalities) ) $coefficients %>% round(., 2)%>% kable() %>% kable_styling()
summary( lm(beertax ~ jail, data = Fatalities) )$coefficients %>% round(., 2)%>% kable() %>% kable_styling()
summary( lm(youngdrivers ~ jail, data = Fatalities) )$coefficients %>% round(., 2) %>% kable() %>% kable_styling()
summary( lm(unemp ~ jail, data = Fatalities) )$coefficients %>% round(., 2) %>% kable() %>% kable_styling()
```

Although covariates `beertax`, `youngdrives`, and `miles` do not show imbalance between treatment (jail = yes) and control (jail = no ) groups, covariates `income` and `unemp` shows noticeable imbalance (p < 0.01 for both), and therefore entails balance adjustments as they will lead to selection bias if unattended.





```{r}
mod1 = glm(jail ~ unemp + income + beertax + youngdrivers + miles,
           family="binomial", data=Fatalities)
summary(mod1)$coef %>% round(., 3) %>% kable() %>% kable_styling()
```
As expected from descriptive analysis, we found that covariate `income` significantly impact the assignment to treatment or control group at p < 0.05. Note that the significance level of `unemp` does not reach p = 0.1, probably due to its strong negative correlation with `income` (r = -0.55) and therefore does not contribute more information when `income` is included in the model. We decide to keep all 5 variables in the selection model regardless of their significance level, as advocated by Caliendo and Kopeinig (2008).


```{r}
prob1 = mod1$fitted.values
pscore1 = ifelse( Fatalities$jail == "yes", prob1,(1-prob1) )
weight1 = 1/pscore1

mod1_new = lmer(frate ~ jail + rescale(unemp) + rescale(income) + rescale(beertax) + rescale(youngdrivers) + rescale(miles) + (1|state), data=Fatalities, weights=weight1)

summary(mod1_new)$coef %>% round(., 3) %>% kable() %>% kable_styling()
```
We used logistic regression to determine the probability of treatment/control assignment, given the 5 chose covariates mentioned earlier. The estimated probablities were then converted to propensity scores, whose inverse are included as weights in the linear mixed model.
After applying propensity score weighting, we still do not found significant treatment effect, as the factor level jailyes is not statistically significant at p = 0.05. Again, the covariates `unemp` and `youngdrivers` are statistically significant at p < 0.001.




```{r}
# balance analysis after propensity score estimation
summary( lm(miles ~ jail, weight = weight1, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(income ~ jail, weight = weight1,data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(beertax ~ jail, weight = weight1,data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(youngdrivers ~ jail, weight = weight1, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(unemp ~ jail, weight = weight1, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
```




## Model 2 (control: breath)



```{r}
mod2 = glm( breath ~  unemp +  income + beertax + youngdrivers + miles, family="binomial", data=Fatalities)
summary(mod2)
```


prob2 = mod2$fitted.values
pscore2 = ifelse(Fatalities$breath == "yes", prob2,(1-prob2))
weight2 = 1/pscore2
mod2_new = lmer( frate ~ breath+rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+rescale(miles)+(1|state),data=Fatalities, weights=weight2)
summary(mod2_new)
```






Model 3 (control: service)
```{r}
mod3 = glm(service ~ rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+rescale(miles),
           family="binomial", data=Fatalities)
prob3 = mod3$fitted.values
pscore3 = ifelse(Fatalities$service == "yes", prob3,(1-prob3))
weight3 = 1/pscore3
mod3_new = lmer(frate~service+rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+
                  rescale(miles)+(1|state),data=Fatalities, weights=weight3)
summary(mod3_new)
```




<!-- Model 4 (control: service) including jail in model -->
<!-- ```{r} -->
<!-- mod4 = glm(service ~ jail+rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+rescale(miles), family="binomial", data=Fatalities) -->
<!-- prob4 = mod4$fitted.values -->
<!-- pscore4 = ifelse(Fatalities$service == "yes", prob4,(1-prob4)) -->
<!-- weight4 = 1/pscore4 -->
<!-- mod4_new = lmer(frate~service+jail+rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+rescale(miles)+(1|year),data=Fatalities, weights=weight4) -->
<!-- summary(mod4_new) -->
<!-- ``` -->



<!-- Model 5 (control: jail) including service in model -->
<!-- ```{r} -->
<!-- mod5 = glm(jail ~ service+rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+ -->
<!--              rescale(miles), family="binomial", data=Fatalities) -->
<!-- prob5 = mod5$fitted.values -->
<!-- pscore5 = ifelse(Fatalities$service == "yes", prob5,(1-prob5)) -->
<!-- weight5 = 1/pscore5 -->
<!-- mod5_new = lmer(frate~service+jail+rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+ -->
<!--                   rescale(miles)+(1|year),data=Fatalities, weights=weight5) -->
<!-- summary(mod5_new) -->
<!-- ``` -->

































==========================================================================================
                    Code used by TA in the discussion session
==========================================================================================


```{r}
dat = Fatalities[, c('fatal', 'beertax', 'jail', 'unemp', 'state', 'youngdrivers')]
dat = dat[!is.na(dat$jail), ]

model1 = lmer(fatal~beertax+jail+unemp+youngdrivers+(1|state),data=dat)
summary(model1)
```
Can we say that jail cannot cause the reduction or increase of fatalities?
We cannot get causal effect from an observational study directly because there exists selection bias

Balance analysis
```{r}
model2 = lm(unemp~jail, data=dat)
summary(model2)
```

To remove the selection bias, we can use propensity scores as weights in our model
First, we need to build a selection model for estimating the propensity scores
The propensity score is defined as the conditional probability given background variables
```{r}
Fatalities$service = relevel(Fatalities$service, ref="no")
models = glm(service ~ beertax+unemp+youngdrivers, family="binomial", data=Fatalities)
prob = models$fitted.values
pscore = ifelse(Fatalities$service == "yes", prob,(1-prob))
weight = 1/pscore

model_new = lmer(frate~beertax+service+unemp+youngdrivers+(1|state),data=Fatalities, weights=weight)
summary(model_new)
anova(model_new)
```




