---
title: "Project 3 Description"
date: "`r Sys.Date()`"
output: html_document
---

## Overview

This document contains instructions on Project 3 for STA 207 in Winter 2021. This document is made with `R markdown`. The `rmd` file to generate this document is available on the course website. 

# Background



## WHO COVID-19 data

We will do an EDA for the WHO COVID-19 data. You can take a look at the [weekly WHO COVID-19 update](https://www.who.int/docs/default-source/coronaviruse/situation-reports/20200921-weekly-epi-update-6.pdf?sfvrsn=d9cf9496_6&download=true) for reference.


```{r echo=TRUE, eval=TRUE}
library(tidyverse)
library(dplyr)
library(magrittr)
library(gridExtra)
covid <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")
```

```{r echo=TRUE, eval=TRUE}
summary(covid)
```

```{r echo=TRUE, eval=TRUE}
# https://ourworldindata.org/covid-face-coverings

# facial_coverings:
# 0: No policy
# 1: Recommended
# 2: Required in some specified shared/public spaces outside the home with other people present, or some situations when social distancing not possible
# 3: Required in all shared/public spaces outside the home with other people present or all situations when social distancing not possible
# 4: Required outside the home at all times regardless of location or presence of other people


face_covering <- read.csv("face-covering-policies-covid.csv")
face_covering$Date <- as.Date(face_covering$Date)
summary(face_covering)
```

```{r echo=TRUE, eval=TRUE}
# covid: Date ranges from 2020-01-03 to 2021-03-03
# face_covering: Date ranges from 2020-01-01 to 2021-03-02

# Subset so the datasets contain date ranging from 2020-01-03 to 2021-03-02

covid %<>% filter (Date_reported >= as.Date("2020-01-03") & Date_reported <= as.Date("2021-03-02"))
face_covering %<>% filter (Date >= as.Date("2020-01-03") & Date <= as.Date("2021-03-02"))
```


```{r echo=TRUE, eval=TRUE}
# Change country names to have the same values from country names in face_covering
covid$Country[covid$Country == "Republic of Korea"] <- "South Korea"
covid$Country[covid$Country == "Russian Federation"] <- "Russia"
covid$Country[covid$Country == "The United Kingdom"] <- "United Kingdom"
covid$Country[covid$Country == "United States of America"] <- "United States"
```


```{r echo=TRUE, eval=TRUE}
# GDP of 25 countries
# https://www.investopedia.com/insights/worlds-top-economies/


covid_GDP25_countries <- covid[grepl("(United States|China|Japan|Germany|India|United Kingdom|France|Italy|Brazil|Canada|Russia|South Korea|Australia|Spain|Mexico|Indonesia|Netherlands|Saudi Arabia|Turkey|Switzerland|Poland|Thailand|Sweden|Belgium|Nigeria)", covid$Country),]

covid_GDP25_countries %<>% filter (Country != "United States Virgin Islands") # drop United States Virgin Islands


face_covering_GDP25_countries <- face_covering[grepl("(United States|China|Japan|Germany|India|United Kingdom|France|Italy|Brazil|Canada|Russia|South Korea|Australia|Spain|Mexico|Indonesia|Netherlands|Saudi Arabia|Turkey|Switzerland|Poland|Thailand|Sweden|Belgium|Nigeria)", face_covering$Entity),]

face_covering_GDP25_countries %<>% filter (Entity != "United States Virgin Islands") # drop United States Virgin Islands
```




```{r echo=TRUE, eval=TRUE}
# The total sum of facial_coverings by country 
# The lower the value, the less restriction (or none) the country has

aggregate(face_covering_GDP25_countries$facial_coverings, by=list(Countries=face_covering_GDP25_countries$Entity), FUN=sum) 
```

```{r echo=TRUE, eval=TRUE, message=FALSE}
# The total number of days that each country didn't have restrictions on facial covering
# Notice that Sweden didn't have facial covering related policy for 370 days


face_covering_GDP25_countries$facial_coverings <- as.factor(face_covering_GDP25_countries$facial_coverings)

country_by_policy <- face_covering_GDP25_countries %>%
    group_by(Entity, facial_coverings) %>%
    summarise(count=n())

country_by_policy[country_by_policy$facial_coverings=="0", ]
```



```{r echo=TRUE, eval=TRUE}
# Not all countries have the same number of observations in the face-covering dataset

face_covering_GDP25_countries %>%
    group_by(Entity) %>%
    summarise(count=n())

```


```{r echo=TRUE, eval=TRUE}
# Whereas, all the countries from the covid dataset has 425 observations. 

covid_GDP25_countries %>%
    group_by(Country) %>%
    summarise(count=n())

```




```{r echo=TRUE, eval=TRUE}
# Left join the two datasets based on countries and date
joined_data <- left_join(covid_GDP25_countries, face_covering_GDP25_countries, by = c("Country" = "Entity", "Date_reported" = "Date"))


# Drop "Code" variable as it is unnecessary to keep
joined_data <- subset(joined_data, select = -Code)

```






```{r echo=TRUE, eval=TRUE}
table(joined_data$WHO_region) / length(unique(joined_data$Date_reported))

# How many countries belong to each WHO region

# EMRO: Eastern Mediterranean Region
# EURO: European Region
# AFRO: African Region
# WPRO: Western Pacific Region
# AMRO: Americans Region
# SEARO: South-East Asian Region
                                                          
```


New_cases
```{r echo=TRUE, eval=TRUE}
new_cases_neg <- which(joined_data$New_cases < 0)

new_cases_neg_matrix <- vector("list",5)
for (i in 1:length(which(joined_data$New_cases < 0))) {
    new_cases_neg_matrix[[i]] <-joined_data[c(new_cases_neg[i]-1, new_cases_neg[i], new_cases_neg[i]+1), c(1,3,5,6)]
}

do.call(rbind,new_cases_neg_matrix)

# When the New_cases has negative values, the Cumulative_cases decreases by that amount (Compare the previous/next day)
```

New_deaths
```{r echo=TRUE, eval=TRUE}
new_deaths_neg <- which(joined_data$New_deaths < 0)

new_deaths_neg_matrix <- vector("list",5)
for (i in 1:length(which(joined_data$New_deaths < 0))) {
    new_deaths_neg_matrix[[i]] <-joined_data[c(new_deaths_neg[i]-1, new_deaths_neg[i], new_deaths_neg[i]+1), c(1,3,7,8)]
}

do.call(rbind,new_deaths_neg_matrix)

# When the New_deaths has negative values, the Cumulative_deaths decreases by that amount (Compare the previous/next day)
```




```{r echo=TRUE, eval=TRUE}
Australia_new_cases <- joined_data %>% filter( Country=="Australia" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Australia") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Belgium_new_cases <- joined_data %>% filter( Country=="Belgium" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Belgium") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Brazil_new_cases <- joined_data %>% filter( Country=="Brazil" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Brazil") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Canada_new_cases <- joined_data %>% filter( Country=="Canada" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Canada") + ylab("") +
  scale_x_date(date_labels = "%b %d")

China_new_cases <- joined_data %>% filter( Country=="China" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("China") + ylab("") +
  scale_x_date(date_labels = "%b %d")

France_new_cases <- joined_data %>% filter( Country=="France" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("France") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Germany_new_cases <- joined_data %>% filter( Country=="Germany" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Germany") + ylab("") +
  scale_x_date(date_labels = "%b %d")

India_new_cases <- joined_data %>% filter( Country=="India" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("India") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Indonesia_new_cases <- joined_data %>% filter( Country=="Indonesia" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Indonesia") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Italy_new_cases <- joined_data %>% filter( Country=="Italy" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Italy") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Japan_new_cases <- joined_data %>% filter( Country=="Japan" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Japan") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Mexico_new_cases <- joined_data %>% filter( Country=="Mexico" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Mexico") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Netherlands_new_cases <- joined_data %>% filter( Country=="Netherlands" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Netherlands") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Nigeria_new_cases <- joined_data %>% filter( Country=="Nigeria" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Nigeria") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Poland_new_cases <- joined_data %>% filter( Country=="Poland" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Poland") + ylab("") +
  scale_x_date(date_labels = "%b %d")

SouthKorea_new_cases <- joined_data %>% filter( Country=="South Korea" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("South Korea") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Russia_new_cases <- joined_data %>% filter( Country=="Russia" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Russia") + ylab("") +
  scale_x_date(date_labels = "%b %d")

SaudiArabia_new_cases <- joined_data %>% filter( Country=="Saudi Arabia" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Saudi Arabia") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Spain_new_cases <- joined_data %>% filter( Country=="Spain" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Spain") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Sweden_new_cases <- joined_data %>% filter( Country=="Sweden" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Sweden") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Switzerland_new_cases <- joined_data %>% filter( Country=="Switzerland" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Switzerland") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Thailand_new_cases <- joined_data %>% filter( Country=="Thailand" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Thailand") + ylab("") +
  scale_x_date(date_labels = "%b %d")

UnitedKingdom_new_cases <- joined_data %>% filter( Country=="United Kingdom" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("United Kingdom") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Turkey_new_cases <- joined_data %>% filter( Country=="Turkey" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("Turkey") + ylab("") +
  scale_x_date(date_labels = "%b %d")

UnitedStates_new_cases <- joined_data %>% filter( Country=="United States" ) %>% 
  ggplot( aes(x=Date_reported, y=New_cases)) + geom_line() + xlab("United States") + ylab("") +
  scale_x_date(date_labels = "%b %d")
```

```{r echo=TRUE, eval=TRUE}
grid.arrange(Australia_new_cases, Belgium_new_cases, Brazil_new_cases, Canada_new_cases, nrow = 2)

grid.arrange(China_new_cases, France_new_cases, Germany_new_cases, India_new_cases, nrow = 2)

grid.arrange(Indonesia_new_cases, Italy_new_cases, Japan_new_cases, Mexico_new_cases, nrow = 2)

grid.arrange(Netherlands_new_cases, Nigeria_new_cases, Poland_new_cases, SouthKorea_new_cases, nrow = 2)

grid.arrange(Russia_new_cases, SaudiArabia_new_cases, Spain_new_cases, Sweden_new_cases, nrow = 2)

grid.arrange(Switzerland_new_cases, Thailand_new_cases, UnitedKingdom_new_cases, Turkey_new_cases, UnitedStates_new_cases,nrow = 3)
```





How face covering-related policy changes over time (need some additional work - see plots below)
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
Australia_facial_coverings <- joined_data %>% filter( Country=="Australia" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Australia") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Belgium_facial_coverings <- joined_data %>% filter( Country=="Belgium" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Belgium") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Brazil_facial_coverings <- joined_data %>% filter( Country=="Brazil" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Brazil") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Canada_facial_coverings <- joined_data %>% filter( Country=="Canada" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Canada") + ylab("") +
  scale_x_date(date_labels = "%b %d")

China_facial_coverings <- joined_data %>% filter( Country=="China" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("China") + ylab("") +
  scale_x_date(date_labels = "%b %d")

France_facial_coverings <- joined_data %>% filter( Country=="France" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("France") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Germany_facial_coverings <- joined_data %>% filter( Country=="Germany" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Germany") + ylab("") +
  scale_x_date(date_labels = "%b %d")

India_facial_coverings <- joined_data %>% filter( Country=="India" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("India") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Indonesia_facial_coverings <- joined_data %>% filter( Country=="Indonesia" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Indonesia") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Italy_facial_coverings <- joined_data %>% filter( Country=="Italy" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Italy") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Japan_facial_coverings <- joined_data %>% filter( Country=="Japan" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Japan") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Mexico_facial_coverings <- joined_data %>% filter( Country=="Mexico" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Mexico") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Netherlands_facial_coverings <- joined_data %>% filter( Country=="Netherlands" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Netherlands") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Nigeria_facial_coverings <- joined_data %>% filter( Country=="Nigeria" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Nigeria") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Poland_facial_coverings <- joined_data %>% filter( Country=="Poland" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Poland") + ylab("") +
  scale_x_date(date_labels = "%b %d")

SouthKorea_facial_coverings <- joined_data %>% filter( Country=="South Korea" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("South Korea") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Russia_facial_coverings <- joined_data %>% filter( Country=="Russia" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Russia") + ylab("") +
  scale_x_date(date_labels = "%b %d")

SaudiArabia_facial_coverings <- joined_data %>% filter( Country=="Saudi Arabia" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Saudi Arabia") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Spain_facial_coverings <- joined_data %>% filter( Country=="Spain" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Spain") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Sweden_facial_coverings <- joined_data %>% filter( Country=="Sweden" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Sweden") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Switzerland_facial_coverings <- joined_data %>% filter( Country=="Switzerland" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Switzerland") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Thailand_facial_coverings <- joined_data %>% filter( Country=="Thailand" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Thailand") + ylab("") +
  scale_x_date(date_labels = "%b %d")

UnitedKingdom_facial_coverings <- joined_data %>% filter( Country=="United Kingdom" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("United Kingdom") + ylab("") +
  scale_x_date(date_labels = "%b %d")

Turkey_facial_coverings <- joined_data %>% filter( Country=="Turkey" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("Turkey") + ylab("") +
  scale_x_date(date_labels = "%b %d")

UnitedStates_facial_coverings <- joined_data %>% filter( Country=="United States" ) %>% 
  ggplot( aes(x=Date_reported, y=facial_coverings)) + geom_line() + xlab("United States") + ylab("") +
  scale_x_date(date_labels = "%b %d")
```


Overlapped periods (e.g. Australia: from mid-November to early-January, two horizontal lines exist) indicate they switched from one policy to another, and back to the original policy 
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
grid.arrange(Australia_facial_coverings, Belgium_facial_coverings, Brazil_facial_coverings, Canada_facial_coverings, nrow = 2)

grid.arrange(China_facial_coverings, France_facial_coverings, Germany_facial_coverings, India_facial_coverings, nrow = 2)

grid.arrange(Indonesia_facial_coverings, Italy_facial_coverings, Japan_facial_coverings, Mexico_facial_coverings, nrow = 2)

grid.arrange(Netherlands_facial_coverings, Nigeria_facial_coverings, Poland_facial_coverings, SouthKorea_facial_coverings, nrow = 2)

grid.arrange(Russia_facial_coverings, SaudiArabia_facial_coverings, Spain_facial_coverings, Sweden_facial_coverings, nrow = 2)

grid.arrange(Switzerland_facial_coverings, Thailand_facial_coverings, UnitedKingdom_facial_coverings, Turkey_facial_coverings, UnitedStates_facial_coverings,nrow = 3)
```


```{r}
arima_function = function(covid_df, country_string, date_cutoff){
  # getting covid data for India
  india.df <- subset( covid, Country == country_string )[, c(1,5) ]
  # fix colnames
  colnames(india.df) <- c( "date", "case" )
  # convert `date` to "Date" object
  india.df$date <- as.Date( india.df$date )
    
  india.ts <- zoo(india.df$case,  india.df$date ) 
  summary(india.ts)
  print(plot(india.ts))
  #policy.start.date <- as.Date( format("2020-04-01", format = "%Y-%m-%d") )
  policy.start.date <- as.Date( format(date_cutoff, format = "%Y-%m-%d") )

  
  # length of days to forecast the number of new cases per day
  forcast.length <-  sum(india.df$date <= policy.start.date)
  
  # length of days when face covering policy is in effect
  # this serves the maximum possible length of dates to forecast
  policy.length <- sum(india.df$date > policy.start.date)
    # subset of data when no face covering
  india.sub.df <- india.df[ india.df$date <= policy.start.date, ]
  # convert to "zoo" object
  india.sub.ts <- zoo(india.sub.df$case,  india.sub.df$date ) 
  
  # subset of data (of length forcast.length) immediate after policy change
  india.real.df <- 
    india.df[ india.df$date > policy.start.date & 
                india.df$date <= (policy.start.date + forcast.length), ]
  
  # subset of data when face covering recommended/required
  india.real.df2 <- india.df[ india.df$date > policy.start.date , ]
  
  
  # ensuring stationary assumption by differencing the data
  # stationary_data <- diff( india.sub.ts )
  stationary_data <- diff ( diff( india.sub.ts ) )
  print(plot( stationary_data ))
  
  # To check if it's stationary we conduct a quantitative test.
  # We use the Augmented Dickey-Fuller Test.
  
  # H_0 = The null hypothesis for this test is that there is a unit root.
  # H_A = The alternative hypothesis is that the time series is 
  # stationary (or trend-stationary).
  adf.test( as.matrix(stationary_data) ) 
  
  # We select a significance level of 0.05 and since our p-value is 0.01
  # and smaller then 0.05, we come to the conclusion to reject the null
  # hypothesis. In other words our data is stationary.
  
  # We use the Auto Correlation Graph
  
  # First we have a look at our acf graph when our data isn't stationary.
  acf( india.sub.ts )
  
  # Here we see that our values are all exceeding the blue line. The goal 
  # is to have the values under the blue line and they should be inverted 
  # as well.
  
  # To select the p and q values we select the number before the first 
  # inverted line.
  acf( stationary_data )
  pacf( stationary_data )
  
  # arima has a auto.arima function which gives us the ideal arima model 
  # based on our data.
  arima_funct <- auto.arima( india.sub.ts )
  arima_funct
  
  
  # the parameter d is 2 from the auto.arima, which means two diff()
  # should be used to transform the `india.sub.ts`
  
  # lets use the auto.arima function to forecast the length of period of
  # no face covering (for India it's 90 days)
  
  forecast1 <- forecast( arima_funct, h = length( india.sub.ts ) )
  forecast1.df <- as.data.frame( forecast1 )
  
  # predicted additional cumulative cases over the next 90 days
  pred_cum_cases <- 
    c( round( sum( forecast1.df$`Lo 95` ), 0 ), 
       round( sum( forecast1.df$`Point Forecast` ), 0),
       round( sum( forecast1.df$`Hi 95` ),0 ) ) 
  
  pred_cum_cases
  
  # actual accumualted cases over the next 90 days
  round( sum( india.real.df$case ), 0 )
  
  # quick-N-dirty way to plot things
  # plot( forecast1, xaxt = "n" )
  # axis(1, at = axTicks(1), labels = as.Date( axTicks(1) ),
  #      cex.axis = 0.7, las = 2 )
  
  
  quick_plot( forecast1, india.sub.df, india.real.df )

  forecast2 <- forecast( object = arima_funct, h = policy.length )
  forecast2.df <- as.data.frame( forecast2 )
  
  # predicted additional cumulative cases over the next 90 days
  pred_cum_cases2 <- 
    c( round( sum( forecast2.df$`Lo 95` ), 0 ), 
       round( sum( forecast2.df$`Point Forecast` ), 0),
       round( sum( forecast2.df$`Hi 95` ),0 ) ) 
  
  pred_cum_cases2
  
  # actual accumualted cases over the next 323 days
  round( sum( india.real.df2$case ), 0 )
  
  quick_plot( forecast2, india.sub.df, india.real.df2 )
}
```





```{r echo=TRUE, eval=TRUE}
countries <- unique(joined_data$Country)
last_date <- 0

for(i in 1:length(countries)) {
  last_date[i] <- (joined_data %>% filter(Country == countries[i]) %>% 
                     filter(facial_coverings != "0" & facial_coverings !="1") %>% head(1))$Date_reported - 1
  
  
}

mask_policy <- zoo::as.Date(last_date)

mask_required <- data.frame(countries, mask_policy)



for(i in 1:3) {
  arima_function(joined_data, mask_required[i,1], mask_required[i,2])
  
  
}



```








```{r echo=TRUE, eval=TRUE}



#missing_countries_face_covering <- c("American Samoa", "Anguilla", "Antigua and Barbuda", "Armenia", "Bonaire", "British Virgin Islands", "Cayman Islands", "Cook Islands", "Curaçao", "Democratic People's Republic of Korea", "Equatorial Guinea", "Falkland Islands (Malvinas)", "French Guiana", "French Polynesia", "Gibraltar", "Grenada", "Guadeloupe", "Guernsey", "Guinea-Bissau", "Holy See", "Isle of Man", "Jersey", "Liechtenstein", "Madagascar", "Maldives", "Marshall Islands", "Martinique", "Mayotte", "Micronesia (Federated States of)", "Montenegro", "Montserrat", "Nauru", "New Caledonia", "Niue", "North Macedonia", "Northern Mariana Islands (Commonwealth of the)", "occupied Palestinian territory, including east Jerusalem", "Other", "Palau", "Pitcairn Islands", "Réunion", "Saba", "Saint Barthélemy", "Saint Helena", "Saint Kitts and Nevis", "Saint Lucia", "Saint Martin", "Saint Pierre and Miquelon", "Saint Vincent and the Grenadines", "Samoa", "Sao Tome and Principe", "Sint Eustatius", "Sint Maarten", "Tokelau", "Turks and Caicos Islands", "Tuvalu", "Wallis and Futuna")


#missing_countries_covid <- c("Hong Kong", "Macao", "Palestine", "Taiwan")
```





This data set is maintained by WHO and updated constantly. The first task for you is to understand this data set, e.g., the meanings of the variables and their values. To this end, you can make use of your preferred search engine, or read the documentation for this data set. 

In this project, you are a team of conscientious statisticians, who wish to help the general public understand the ongoing pandemic. 



## Suggested outline 

The following list provides one potential structure of the data analysis report. As this is the final project, the following suggestions are intended to provide one viable route for your project while leaving you as much freedom as possible. 

***

0. Before writing your analysis report, you may want to explore this data set and read about the coronavirus to generate the hypothesis or question to be answered in this report, i.e., the question(s) of interest. You can be creative on this question so long as it meets two conditions.
    a. This question might be meaningful in real-world. 
    b. A method discussed in this course is employed in solving this question. 

***


1. Introduce the data set and the questions of interest. 

2. Review the background of the coronavirus. 

3. Explore this dataset and generate summary statistics that you find informative, and explain your findings. <span style="color:red"> The summary statistics should include at least time, number of cases, number of death, case-mortality rate.</span>

4.  Propose an appropriate model to answer the questions of interest. 
    a. Explain your notation. 
    b. State assumptions for your model. 
    c. Explain why your model is appropriate for this task. 


5.  Fit the proposed model in (4) and explain your results. 

6.  <span style="color:red"> Discuss whether you can make any causal inference in this analysis. </span>

7.  Conduct model diagnostics and/or sensitivity analysis. 

8. Conclude your analysis with an discussion of your findings and caveats of your approach. 




