---
title: "Project 2"
author: "Seyoung Jung"
date: "2/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r,warning=FALSE,message=FALSE}
library(AER)
library(tidyverse)
library(lubridate)
library(lmerTest)
library(scales)
library(kableExtra)
```


```{r}
data('Fatalities')
```


```{r}
summary(Fatalities)
```

Observe the row(s) containing NA value in "jail" and/or "service".
One row (row #28) has NA in the two variables.
```{r}
Fatalities[which(is.na(Fatalities$jail) | is.na(Fatalities$service)), ]
```


Observe the row which has the maximum value from the "miles" variable
Notice the stark difference between the highest and the second highest average miles per driver 
```{r}
Fatalities[order(-Fatalities$miles), ] %>% head(5)

plot(Fatalities$miles[-171], Fatalities$fatal[-171], xlab = 'Average miles per driver', ylab = 'The number of fatalities' )

plot(Fatalities$miles[-171], Fatalities$frate[-171], xlab = 'Average miles per driver', ylab = 'Fatality rate' )
```


Check whether the other rows of California have data on "jail" and "service".
The other rows have "no"s in the jail and service columns. 
```{r}
Fatalities[which(Fatalities$state=="ca"), ]
```


Convert the "state", "year", "drinkage" into a factor variable
```{r}
Fatalities$drinkage = round(Fatalities$drinkage)
Fatalities$drinkage = as.factor(Fatalities$drinkage)
```


Correlation between "fatal" and the other non-categorical variables
```{r}
cor(Fatalities$fatal, Fatalities[, -which(sapply(Fatalities, class)=="factor")])
```


Create a new variable "frate" which is the total number of fatalities over the population of the state. 
Correlation between "frate" and the other non-categorical variables
Some variables (e.g. income, beertax, baptist, miles, and youngdrivers) have relatively high correlations with "frate".
```{r}
Fatalities$frate = Fatalities$fatal/Fatalities$pop

cor(Fatalities$frate, Fatalities[, -which(sapply(Fatalities, class)=="factor")])
```








```{r}
#summary(glm(breath~frate,data=Fatalities,family=binomial))$coefficients # breath vs. frate
#summary(glm(jail~frate,data=Fatalities,family=binomial))$coefficients # jail vs. frate
#summary(glm(service~frate,data=Fatalities,family=binomial))$coefficients # service vs. frate

#summary(glm(breath~unemp,data=Fatalities,family=binomial))$coefficients # breath vs. unemp
#summary(glm(jail~unemp,data=Fatalities,family=binomial))$coefficients # jail vs. unemp
#summary(glm(service~unemp,data=Fatalities,family=binomial))$coefficients # service vs. unemp

#summary(glm(breath~miles,data=Fatalities,family=binomial))$coefficients # breath vs. miles
#summary(glm(jail~miles,data=Fatalities,family=binomial))$coefficients # jail vs. miles
#summary(glm(service~miles,data=Fatalities,family=binomial))$coefficients # service vs. miles

#summary(glm(breath~youngdrivers,data=Fatalities,family=binomial))$coefficients # breath vs. youngdrivers
#summary(glm(jail~youngdrivers,data=Fatalities,family=binomial))$coefficients # jail vs. youngdrivers
#summary(glm(service~youngdrivers,data=Fatalities,family=binomial))$coefficients # service vs. youngdrivers

#summary(glm(breath~beertax,data=Fatalities,family=binomial))$coefficients # breath vs. beertax
#summary(glm(jail~beertax,data=Fatalities,family=binomial))$coefficients # jail vs. beertax
#summary(glm(service~beertax,data=Fatalities,family=binomial))$coefficients # service vs. beertax

#summary(glm(breath~income,data=Fatalities,family=binomial))$coefficients # breath vs. income
#summary(glm(jail~income,data=Fatalities,family=binomial))$coefficients # jail vs. income
#summary(glm(service~income,data=Fatalities,family=binomial))$coefficients # service vs. income
```


```{r}
# explore  relationships between jail and unemployment
ggplot(data=Fatalities, aes(unemp, fill = jail)) + 
          geom_density( alpha = 0.5 )
```
unemployment does not seem to strongly impact the propensity of being assigned to treatment (jail = yes) or control (jail = no) group.

```{r}
# explore relationships between jail and income
ggplot(data=Fatalities, aes(income, fill = jail)) + 
          geom_density( alpha = 0.5 )

```
It seems that income impacts the propensity of being assigned to treatment (jail = yes) or control (jail = no) group. Therefore income should be included in the selection model.

```{r}
# explore relationships between jail and fatalities
ggplot(data=Fatalities, aes(miles, fill = jail)) + 
          geom_density( alpha = 0.5 )

```
mileage driven per person does not seem to strongly impact the propensity of being assigned to treatment (jail = yes) or control (jail = no) group.

```{r}
# explore relationships between jail and beertax
ggplot(data=Fatalities, aes(beertax, fill = jail)) + 
          geom_density( alpha = 0.5 )
```

beertax does not seem to strongly impact the propensity of being assigned to treatment (jail = yes) or control (jail = no) group.

```{r}
# explore relationships between jail and youngdrivers
ggplot(data=Fatalities, aes(youngdrivers, fill = jail)) + 
          geom_density( alpha = 0.5 )
```

percent of young drivers does not seem to strongly impact the propensity of being assigned to treatment (jail = yes) or control (jail = no) group.

```{r}
# Adpated from http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
my_cols <- c("#00AFBB", "#FC4E07")  

# Correlation panel
panel.cor <- function(x, y){
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y)
    txt <- paste0("r = ", round(r, digits = 2))
    text(0.5, 0.5, txt, cex = 3*abs(r) )
}

# Customize upper panel
upper.panel<-function(x, y){
  points(x,y, pch = 19, col = alpha( my_cols[Fatalities$jail], 0.3 ) )
}

```

```{r}
pairs( Fatalities[, c('income', 'unemp', 'miles', 'beertax', 'youngdrivers') ],
       upper.panel = upper.panel,
       lower.panel = panel.cor )
```



==========================================================================================
                              Model Building (Seyoung Jung)
==========================================================================================

```{r}
 cor(Fatalities$frate, Fatalities[, -which(sapply(Fatalities, class)=="factor")])  


```
Independent variables chosen based on the degree of correlation with "frate"

We start by building a tentative outcome model without the use of propensity scores.

## Model 1 (treatment (jail = yes) and control (jail = no))

```{r warning=FALSE}

summary( lmer(frate ~ jail + rescale(unemp) + rescale(income) + rescale(beertax) + rescale(youngdrivers) +  rescale(miles) + (1|state), data=Fatalities) )$coef %>% round(.,3) %>%  kable() %>% kable_styling()

```

We found a non-significant treatment effect, as the factor level jailyes is not statistically significant at p = 0.05. However, the covariates `unemp` and `youngdrivers` are statistically significant at p < 0.001.

Based on our exploratory data analysis and group discussion, we assume that the variables `unemp`, `income`, `beertax`, `youngdrivers` and `miles` have an influence on the assignment to either the control (jail = no) or the treatment (jail = yes) group.

Next we performed a balance analysis, trying to assess the degree of bias during assignment to treatment and control groups, before the use of propensity scores.

```{r}
# balance analysis before estimating propensity scores
summary( lm(miles ~ jail, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(income ~ jail, data = Fatalities) ) $coefficients %>% round(., 3)%>% kable() %>% kable_styling()
summary( lm(beertax ~ jail, data = Fatalities) )$coefficients %>% round(., 3)%>% kable() %>% kable_styling()
summary( lm(youngdrivers ~ jail, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(unemp ~ jail, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
```

This time it turns out that covariate `unemp` shows strong imbalance between treatment (jail = yes) and control (jail = no ) groups (p < 0.01), and covariate `income` shows a marginal significant imbalance (p < 0.05). Although not significant, the significance levels of covariates `miles` and `beertax` are close to p = 0.05 level, also suggesting potential selection bias. These should be addressed using propensity scores.


```{r}
mod1 = glm(jail ~ unemp + income + beertax + youngdrivers + miles,
           family="binomial", data=Fatalities)
summary(mod1)$coef %>% round(., 3) %>% kable() %>% kable_styling()
```
We found that covariate `unemp` significantly impact the assignment to treatment or control group at p < 0.001, and covariate `miles` significantly impact the assignment to treatment or control group at p < 0.01. Again, all 5 variables are included in the selection model regardless of their significance level,


```{r}
prob1 = mod1$fitted.values
pscore1 = ifelse( Fatalities$jail == "yes", prob1,(1-prob1) )
weight1 = 1/pscore1

mod1_new = lmer(frate ~ jail + rescale(unemp) + rescale(income) + rescale(beertax) + rescale(youngdrivers) + rescale(miles) + (1|state), data=Fatalities, weights=weight1)

summary(mod1_new)$coef %>% round(., 3) %>% kable() %>% kable_styling()
```
We used logistic regression to determine the probability of treatment/control assignment, given the 5 chose covariates mentioned earlier. The estimated probablities were then converted to propensity scores, whose inverse are included as weights in the linear mixed model.
After applying propensity score weighting, we still do not found significant treatment effect, as the factor level jailyes is not statistically significant at p = 0.05. Again, the covariates `unemp` and `youngdrivers` are statistically significant at p < 0.001.


```{r}
# balance analysis after propensity score estimation
summary( lm(miles ~ jail, weight = weight1, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(income ~ jail, weight = weight1,data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(beertax ~ jail, weight = weight1,data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(youngdrivers ~ jail, weight = weight1, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(unemp ~ jail, weight = weight1, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
```


The selection biases by `income` and `unemp` are reduced much after propensity score weighting.


## Model 2 (control: breath)

```{r warning=FALSE}

summary( lmer(frate ~ breath + rescale(unemp) + rescale(income) + rescale(beertax) + rescale(youngdrivers) +  rescale(miles) + (1|state), data=Fatalities) )$coef %>% round(.,3) %>%  kable() %>% kable_styling()

```

We found a marginally treatment effect, as the factor level breathyes is statistically significant at p = 0.05. Interestingly, the covariates `unemp` and `youngdrivers` are statistically significant at p < 0.001.


Likewise, we performed a balance analysis before the use of propensity scores.

```{r}
# balance analysis before estimating propensity scores
summary( lm(miles ~ breath, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(income ~ breath, data = Fatalities) ) $coefficients %>% round(., 3)%>% kable() %>% kable_styling()
summary( lm(beertax ~ breath, data = Fatalities) )$coefficients %>% round(., 3)%>% kable() %>% kable_styling()
summary( lm(youngdrivers ~ breath, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(unemp ~ breath, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
```

This time it turns out that covariate `unemp` shows strong imbalance between treatment (jail = yes) and control (jail = no ) groups (p < 0.01), and covariate `income` shows a marginal significant imbalance (p < 0.05). Although not significant, the significance levels of covariates `miles` and `beertax` are close to p = 0.05 level, also suggesting potential selection bias. These should be addressed using propensity scores.


```{r}
mod2 = glm(breath ~ unemp + income + beertax + youngdrivers + miles,
           family="binomial", data=Fatalities)
summary(mod2)$coef %>% round(., 3) %>% kable() %>% kable_styling()
```
 We found that covariate `unemp` significantly impact the assignment to treatment or control group at p < 0.001, and covariate `miles` significantly impact the assignment to treatment or control group at p < 0.0. Again, all 5 variables are included in the selection model regardless of their significance level.


```{r}
prob2 = mod2$fitted.values
pscore2 = ifelse(Fatalities$breath == "yes", prob2,(1-prob2))
weight2 = 1/pscore2
mod2_new = lmer( frate ~ breath+rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+rescale(miles)+(1|state),data=Fatalities, weights=weight2)
summary(mod2_new)$coef %>% round(., 3) %>% kable() %>% kable_styling()
```
Likewise, after applying propensity score weighting, we still do not found significant treatment effect, as the factor level breathyes is not statistically significant at p = 0.05. Nevertheless, it is significant at p < 0.1, indicating potential effects that needs further investigation. Again, the covariates unemp and youngdrivers are statistically significant at p < 0.001.

```{r}
# balance analysis after propensity score estimation
summary( lm(miles ~ breath, weight = weight2, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(income ~ breath, weight = weight2,data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(beertax ~ breath, weight = weight2,data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(youngdrivers ~ breath, weight = weight2, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(unemp ~ breath, weight = weight2, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
```

Note that the selection biases by `income` and `unemp` are not are reduced much after propensity score weighting, so there still exists substantial selection biases that may invalidate causal inferences.


Model 3 (control: service)
```{r}
mod3 = glm(service ~ rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+rescale(miles),
           family="binomial", data=Fatalities)
prob3 = mod3$fitted.values
pscore3 = ifelse(Fatalities$service == "yes", prob3,(1-prob3))
weight3 = 1/pscore3
mod3_new = lmer(frate~service+rescale(unemp)+rescale(income)+rescale(beertax)+rescale(youngdrivers)+
                  rescale(miles)+(1|state),data=Fatalities, weights=weight3)
summary(mod3_new)$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
```

Using the same method, we do not found significant treatment effect, as the factor level serviceyes is not statistically significant at p = 0.05 Again, the covariates `unemp` and `youngdrivers` are statistically significant at p < 0.001.

```{r}
# balance analysis before propensity score estimation
summary( lm(miles ~ service, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(income ~ service,data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(beertax ~ service, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(youngdrivers ~ service, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(unemp ~ service, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()

# balance analysis after propensity score estimation
summary( lm(miles ~ service, weight = weight3, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(income ~ service, weight = weight3,data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(beertax ~ service, weight = weight3,data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(youngdrivers ~ service, weight = weight3, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
summary( lm(unemp ~ service, weight = weight3, data = Fatalities) )$coefficients %>% round(., 3) %>% kable() %>% kable_styling()
```

It seems that the selection biases by `beertax` are effectively removed after propensity score weighting.
